Refining from the VHDL-AMS Hardware Description Language to the BON High-level Specification Language

### 2010-01-06 papers/BON_VHDL/paper.tex
Rewrote another section or two.

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\ent}{entity\xspace}
\newcommand{\arch}{architecture\xspace}
\newcommand{\archs}{architectures\xspace}
\newcommand{\inv}{\emph{invariant}\xspace}
\newcommand{\undef}{\emph{undefined}\xspace}
\newcommand{\arcoss}{\emph{across}\xspace}
\newcommand{\through}{\emph{through}\xspace}

@@

Consequently, we have developed an approach through which we abstract
(via a formal refinement) a VHDL-AMS specification up to a formal BON
specification.  Using this specification, we then generate (again, via
a formally defined refinement) an executable, functional, model-based
specification in the JML language.  We have subsequently used this
specification to implement a fully validated and verified \ST
simulator.  Due to the semantic properties of these refinements, as
seen in~\autoref{diagram}, we are guaranteed that all functional
properties of the original VHDL-AMS specification are preserved in the
software simulator, thereby witnessing full hardware-software
co-design in a formal sense.

XXX

The \ST sensor platform, represents heterogeneous collection of
digital, analog and mixed signal hardware circuits.  The complexity of
our platform requires the need to strive for formal descriptions each
component separately and behavior of their complex interdependencies,
in order to produce executable specification.\footnote{This is not a
  good reason for performing this research.  Complexity is often the
  reason to use \emph{less} formality, nor more.  Also, why must the
  specification be \emph{executable}?}  To make sure that the sensor
platform works properly we need to translate the VHDL-AMS code into
the specification, and to refine and verify this
specification.\footnote{Too vague a statement---what are the
  \emph{real} reasons we are doing this work?}  In other words, our
aim is to establish formal link between behavior of a given sensor
platform, and derived specification of components in order to
implement fully validated and verified \ST simulator.  For this
purpose, we need notation and methodology which will be able to
satisfy the requirements given below.

First, both of them, should be in sync with VHDL-AMS language and
reference manual which tries to define VHDL-AMS as well as possible in
a descriptive way, explaining the semantics in English.

Second, specification derived from the notation should be easy to
attach\footnote{Meaning?} to the rest of the system
architecture\xspace(driver, microprocessor emulator, test-benches
e.t.c.).

Third, methodology must aid efforts on creation of target system
correctness criteria, such that are independent of implementation.  It
means that verification of simulator model should take place on
specification using reasoning techniques provided for algebraic
formalism before coding starts.

Fourth, a precise semantics should provides the basis for automated
forward-backward engineering.

Managing these requirements, and insuring that the platform definition
(constraints) are preserved and verified through all levels of
abstraction is where one becomes aware of gaps in already used design
methodologies.  In our opinion, the BON-JML approach is in line with
requirements given above.  The authors will illustrate in this article
how BON can represent a high level analogue mapping notation and will
further illustrate a refinement technique from VHDL-AMS to BON and
JML.

This article provides three main contributions.
\begin{enumerate}
\item It provides a solution how to formally emulate piece of hardware
using BON-JML approach; up to date BON and JML have been used mostly
in the pure object oriented software construction.
\item It relates hardware description code and refinement technique.
It means that if a satisfactory VHDL-AMS code exists, then it can be
simulated from the BON (informal/formal) specification, and JML
implementation by an appropriate sequence of refinements.
\item It introduces \inv as a general concept of modeling hardware
components behavior and their interactions.  To be more precisely, the
behavioral specification of the \STS may be represented as a \inv on
the signals at the circuit's ports and may be obtained by taking the
conjunction of the \emph{invariants} that describe behaviors of the
individual components separately.

\end{enumerate} The rest of this article is organized as follows.  In
\autoref{sec:vhdlams} we have presented an overview of VHDL-AMS
language.  \autoref{sec:choiceobon} defines BON principles and
explains the reasons why we have decided to use BON method.  In
\autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
The central part of the our article is \autoref{sec:refVB}, in which
we will give practical aspect of refinement, applied on first order
passive RC low-pass filter, in order to produce formal specification
and software simulation. In this section we explore this approach in
detail.  In \autoref{sec:conclusion} the conclusions are given, and
directions of future studies are outlined. Finally, a comprehensive
list of the mapping between BON and VHDL-AMS can be found in the
appendix.

\section{VHDL-AMS}
\label{sec:vhdlams} IEEE VHDL-AMS 1076.1 is the industry standard
mixed-signal high-level description language for electronic and
multi-domain systems. VHDL-AMS, as an superset of VHDL, supports
modeling and simulation of analog and mixed-signal systems. The main
contribution of VHDL-AMS is ability to describe the behavior of
complex system trought differential and algebraic equations
(DAEs). For this purpose, VHDL-AMS introduces the following
capabilities and therms:

\begin{itemize}
\item \emph{Terminals} represents phisical connection points or
circuit nodes of a network connectivity. They are declared to be of
various \emph{natures}, which represent different energy domains of a
system.  Terminals hold no values; their only role is to facilitate
the formation of conservative equation sets that, in turn, constrain
the associated \emph{branch qualities}.
\item The \emph{quantities} denote a waveform or a time series of
values.  The behavior of the \emph{quantities} is given by the
\emph{simultaneous statements}. They take values as a result of
solving the set of simultaneous ordinary differential and algebraic
equations (DAEs). To simplify, quantities can be understood as analog
unknowns of (DAEs) that specify analog behaviour. There are three
kinds of quantity in VHDL-AMS: free, branch and source quantities. A
\emph{free quantity} can be used in signal-flow modeling. A
\emph{branch quantity} is used to model conservative energy systems. A
\emph{source quantity} is used for frequency and noise modeling.
\item The \emph{simultaneous statements} creates one or more
characteristic equations that are solved by the dedicated simulation
kernel a.k.a.  "analog solver", to model continuously valued behavior.
\item The \emph{procedural statements} or \emph{procedural} express
analog behaviour using sequential statements. Actually procedural
describes analog behaviour by computing new values for quantities from
the current value of quantities.
\item The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient simulation
and to mark discontinuities during such a transient simulation. For
any concurrent break statement there is an associated process.  A
break statement allows restart of the analog kernel, than overrides
quantity values and executes concurrently with the analogue model.

\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon} %Brief description and motivation.  The
Business Object Notation (BON) is an analysis and design notation for
object-oriented systems which brings \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.  Let us define
these principles precisely and point out why the principles are
important in VHDL-AMS simulation.

\textbf{Seamlessness and Reversibility} \xspace \xspace Seamlessness
is the use of a continuous process throughout the software
life-cycle. Reversibility is the support for both forward and backward
development process: from analysis to design and implementation, and
back.  In VHDL-AMS, modelling represents an continuous
top-down-bottom-up process. Top-down approach serves for design while
bottom-up serves for verification. The possibility of seamlessly
translating changes made during a certain development phase back into
earlier phases, facilitates a stable and predictable environment and
ensures the various models consistent with the code. Keeping in sync
various models, documentation and specification from forward progress
to backward and redesign, preclude that design and implementation
diverge to an inconsistent state.

\emph{Seamlessness} also plays an important role in relating behaviors
at different levels of abstraction. For instance, one may wish to
analyze a "lowpass-filter" at the level of individual components (a
resistor and capacitor) and their behaviours, or at the more abstract
level through general behaviour (\inv).  Therefore, capturing
similarities between the differing levels of abstractions, making it
possible to obtain a continuous process that ensures a direct mapping
between a formal specification and its software solution.

\textbf{Contracting} \xspace \xspace Contract specification describe
system components interaction, based on mutual obligations and
benefits, and, for each software element, associated semantic
properties. BON use a full assertion mechanism, allowing analysts to
specify both:\xspace (1) the \emph{structure} of a system and
(2)\xspace \emph{semantics} ( constraints and invariants).  The common
way to specify a program statement is to give a \emph{precondition}
and \emph{postcondition} on the state that the implementation must
satisfy, assuming that only certain program variables are changed.  In
VHDL-AMS any predefined (implicate) attribute has defined \emph{pre}
(\emph{post}) \emph{condition} trought input (output) parameters and
theirs constraints. Furthermore, any subtype (subnature) represents
restricted set of values from a base type (nature).  The condition
that determines which values are in the subtype (subnature) may be
contracted by specifying characteristic predicate which represents
\inv.

Let us next look how \emph{simultaneous statements} are handled in
contract design.  The \emph{simultaneous statement} represents
analogue constraint of continuous-time waveform. We will assume that
the constraint of waveform represents \inv. This premise deserve an
informal explanation.

Briefly stated, analog kernel calculate quantities such that the
constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations. It means that the number of simultaneous
statements should be equal to the number of unknowns
(\emph{quantities}).  For instance, when we reasoning about the
conservation of energy law, the \emph{simultaneous statements}
constrain the values of \emph{branch quantities} in a model.  In other
words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across
quatities =0}) around any closed path must be equal to zero and the
sum of \through quantities (\texttt{$\sum_{node}$ through quatities
=0}) at any node (terminal) must be equal to zero. To be more precise,
the results of summations above remains zero when a particular
transformation is applied to quantities.  In general, they are \inv.

@@

Therefore, on the basis of the above, BON is the obvious choice for
the \STS.  Throughput this article we will make this more clearly.
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB} In this section, we briefly describe the mapping
between VHDL-AMS and BON notation which is discussed in the points
below.

@@

\subsection{Low-Pass filter transfer function }
\label{subsec:lptransfunc} 
\input{lowpassRC}

The most simple low-pass filter is passive RC network shown in
\autoref{fig:lowpassfig}.  According to equation given above (see
\autoref{fig:transfun}) we can produce VHDL-AMS \emph{simultaneous
  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
transformation looks as follows:
\begin{equation}
v_{in} = v_{out} + \tau_psv_{out} 
   \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
   \end{array} 
v_{out} + \tau_p\cfrac{dv_{out}}{dt}
\begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
   \end{array}
  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
\end{equation}

The last equation can be described as follows.  The input
voltage\xspace($v_{in}$)\xspace equals summation of output
voltage\xspace($v_{out}$)\xspace and multiplication of the time
constant\xspace($\tau_p$)\xspace of the pole and the first derivative
of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace over time.

Now we can discretize previous equation. Let us represent $v_{in}$ by
a series of samples $v_{in(1..n)}$. At the same points in time, we can
represent $v_{out}$ by series of sample $v_{out(1..n)}$. We can assume
that samples are taken at points in time incrementing by
$\triangle$t.

The last equation gives us a way to calculate the output samples in
terms of the input samples and the preceding output.  But how to
represent these unknowns in BON ?  We introduce the
\lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
with features as follows:
\begin{itemize}
\item
\emph{across} which represent current:$i_{out}$ 
\item 
\emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
\item 
\emph{terminal sequences} which describes input/output terminal(s). 
\end{itemize}
Now we are in presence of BON notation and we may start with the
refinement of the VHDL-AMS code.

@@

%=====================================================================
\section{Conclusion}
\label{sec:conclusion} The BON notation has originally been used for
description of software models. However, after certain elaboration it
can be used for description of hardware components. Practical
utilization of the approach in the project on mapping between VHDL-AMS
and BON using refinement technique, has demonstrated through low-pass
filter example and shown that simulation of component requires
relatively little and inexpensive efforts. Moreover, our article shown
that the development of the specification produced by the BON and
translated into JML can be applied as refinement procedure: from
VHDL-AMS to informal BON, next, from informal to more formal BON and
finally, to JML specification.  The BON support general software
development and should be adapted for concrete applications and
language as VHDL-AMS. For this purpose we have developed various
VHDL-AMS related types, operators and relations and we have defined
appropriate specifications for all of them. However, the behavior of
\texttt{break} statement and concurrency, for example, have not been
translated yet in the BON notation. We have started a new phase of
project named \emph{eBon} in which the issues mentioned above, and
creation of appropriate components and relations are main subject.

>>>>

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\ent}{entity\xspace}
\newcommand{\arch}{architecture\xspace}
\newcommand{\archs}{architectures\xspace}
\newcommand{\inv}{\emph{invariant}\xspace}
\newcommand{\undef}{\emph{undefined}\xspace}
\newcommand{\arcoss}{\emph{across}\xspace}
\newcommand{\through}{\emph{through}\xspace}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}
    #1}}

@@

Consequently, we have developed an approach through which we abstract
(via a formal refinement) a VHDL-AMS specification up to a formal BON
specification.  Using this specification, we then generate (again, via
a formally defined refinement) an executable, functional, model-based
specification in the JML language.  We have subsequently used this
specification to implement a fully validated and verified \ST
simulator.

Due to the semantic properties of these refinements, as seen
in~\autoref{diagram}, we are guaranteed that all functional properties
of the original VHDL-AMS specification are preserved in the software
simulator, thereby witnessing full hardware-software co-design in a
formal sense.

In summary, this article provides three main contributions.
\begin{enumerate}
\item We define a relation between a hardware description language,
  VHDL-AMS, and a high-level specification language (BON).  This
  relation means that, given a well-formed VHDL-AMS program, it can be
  described using a BON (informal/formal) specification, and hence to
  a JML implementation, through a sequence of refinements.
\item We describe how to formally emulate piece of hardware using a
  refinement-based approach.  Our approach uses object-oriented
  languages, including a high-level and a low-level specification
  language, BON and JML respectively.
\item We introduce the use of \emph{invariants} as a general concept
  for modeling hardware components behavior and their interactions.
  To be more precise, the behavioral specification of a hardware
  component is represented as an invariant on the signals at a
  circuit's ports, which in turn is derived from the conjunction of
  the invariants that describe behaviors of the component's
  subsystems.
\end{enumerate}

The rest of this article is organized as follows.  In
\autoref{sec:vhdlams} we present an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines BON principles and explains the
reasons why we have used the BON method.  In \autoref{sec:mappingVB}
the relation between VHDL-AMS and BON is summarized.  A comprehensive
description of the mapping between BON and VHDL-AMS is found in a
separate technical report.  This refinement is applied to a small
example in \autoref{sec:refVB}, in which we abstract from a passive RC
low-pass filter in order to produce formal specification and software
simulation.  In \autoref{sec:conclusion} we reflect on this work and
point to some open problems and research opportunities.

%=====================================================================
\section{VHDL-AMS}
\label{sec:vhdlams}

The VHDL-AMS language, defined by IEEE VHDL-AMS 1076.1, is the
industry standard mixed-signal high-level description language for
electronic and multi-domain systems.  VHDL-AMS, as an superset of
VHDL, supports modeling and simulation of \emph{analog} and
mixed-signal systems.

\subsection{Key Qualities of VHDL-AM}
\label{subsec:key-qualities-vhdl}

The main contribution of VHDL-AMS is ability to describe the behavior
of a complex system through differential and algebraic equations
(DAEs).  For this purpose, VHDL-AMS introduces the following
notions:
\begin{itemize}
\item \emph{Terminals} represents physical connection points or
  circuit nodes of a network connectivity.  Each terminal has a
  \emph{nature}, which represents different energy domains of a
  system.  Terminals hold no values---their only role is to facilitate
  the formation of conservative equation sets that, in turn, constrain
  the associated \emph{branch qualities}.

\item \emph{Quantities} denote a waveform or a time series of values.
  The behavior of the \emph{quantities} is given through the use of
  \emph{simultaneous statements}.  Quantities take values as a result
  of solving a set of simultaneous ordinary differential and algebraic
  equations (DAEs).  \notej{We need a concrete example here.  This
    does not make much sense in isolation.  Perhaps this detail should
    just be moved or replicated later.  -JRK}

  To simplify, quantities are understood as analog unknowns of (DAEs)
  that specify analog behaviour.  There are three kinds of quantities
  in VHDL-AMS: free, branch, and source quantities.  A \emph{free
    quantity} is used in signal-flow modeling.  A \emph{branch
    quantity} is used to model conservative energy systems.  A
  \emph{source quantity} is used for frequency and noise modeling.

\item \emph{Simultaneous statements} define one or more characteristic
  equations that are solved by the dedicated simulation kernel, a.k.a.
  an "analog solver," to model continuously valued behavior.

\item \emph{Procedural statements} express analog behaviour using
  sequential statements. Actually, procedural statements describe
  analog behaviour by computing new values for quantities from the
  current value of quantities.

\item The \emph{break statement} is an construct used to initialize
  the DAE system at the beginning of a transient simulation, and to
  mark discontinuities during such a transient simulation.  For any
  concurrent break statement there is an associated process.  A break
  statement allows restart of the analog kernel, than overrides
  quantity values and executes concurrently with the analogue model.
  \notje{I have no idea what this last sentence means. -JRK}
\end{itemize}

\subsection{A Running Example: A Low-Pass Filter}
\label{subsec:lptransfunc}

\input{lowpassRC}

A simple low-pass filter is a passive RC network shown in
\autoref{fig:lowpassfig}.  According to the equation given above (see
\autoref{fig:transfun}), we can produce VHDL-AMS \emph{simultaneous
  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
transformation looks as follows:
\begin{equation}
v_{in} = v_{out} + \tau_psv_{out} 
   \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
   \end{array} 
v_{out} + \tau_p\cfrac{dv_{out}}{dt}
\begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
   \end{array}
  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
\end{equation}

This last equation is interpreted as follows.  The input voltage
($v_{in}$) equals the sum of output voltage ($v_{out}$) and
multiplication of the time constant ($\tau_p$) of the pole and the
first derivative of the capacitor's voltage ($v_{out}{'dot}$) over
time.

Now we discretize previous equation.  Let us represent $v_{in}$ by a
series of samples $v_{in(1..n)}$.  At the same points in time, we
represent $v_{out}$ by a series of samples $v_{out(1..n)}$.  We assume
that samples are taken at points in time incrementing by $\triangle$t.

The last equation gives us a way to calculate the output samples in
terms of the input samples and the preceding output.

But how to represent these unknowns in BON\notej{This whole last bit
  should be moved to after we introduce BON and the VHDL-AMS
  metamodel.}?  We introduce the \lstinlinen|BRANCH_QUANTITY| objects
like wrappers around unknowns, with features as follows:
\begin{itemize}
\item \emph{across} which represent current:$i_{out}$ 
\item \emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
\item \emph{terminal sequences} which describes input/output
  terminal(s). 
\end{itemize}

%===================================================================== 
\section{The Business Object Notation}
\label{sec:choiceobon}

The Business Object Notation (BON) is an analysis and design notation
for object-oriented systems which supports \emph{seamlessness},
\emph{reversibility}, and software \emph{contracting}.  Let us define
these principles precisely and point out why they are important in
VHDL-AMS simulation.

\paragraph{Seamlessness and Reversibility} \emph{Seamlessness} is the
use of a continuous process throughout the software life-cycle.
\emph{Reversibility} is the support for both forward and backward
development process: from analysis to design and implementation, and
back.

In VHDL-AMS, modeling represents an continuous top-down/bottom-up
process.  A top-down approach supports design, while bottom-up
supports verification.  The possibility of seamlessly translating
changes made during a certain development phase back into earlier
phases facilitates a stable and predictable environment and ensures
the various models consistent with the code.  Keeping in sync various
models, documentation and specification from forward progress to
backward and redesign preclude that design and implementation diverge
to an inconsistent state.

\emph{Seamlessness} also plays an important role in relating behaviors
at different levels of abstraction.  For instance, one may wish to
analyze a lowpass-filter at the level of individual components (a
resistor and capacitor) and their behaviours or, at the more abstract
level, through general behaviour expressed via an invariant.
Therefore, capturing similarities between the differing levels of
abstractions makes it possible to obtain a continuous process that
ensures a direct mapping between a formal specification and its
software solution.

\paragraph{Contracting} Contract specifications describe system
components and their interactions based on mutual obligations, and,
for each software element, its associated semantic properties.  BON
use an assertion mechanism, supporting analysts' need to specify
both:~(1) the \emph{structure} of a system, and (2)~the
\emph{semantics} of a system via a set of constraints and invariants.

The common way to specify a program statement is to give a
\emph{precondition} and \emph{postcondition} on the state that the
implementation must satisfy, assuming that only certain program
variables are changed.  In VHDL-AMS, any predefined (implicate)
attribute has defined \emph{pre} (\emph{post}) \emph{condition}
through input (output) parameters and their constraints.  Furthermore,
any subtype (subnature) represents restricted set of values from a
base type (nature).  The condition that determines which values are in
the subtype (subnature) may be contracted by specifying a
characteristic predicate which represents an invariant.

Let us next look how \emph{simultaneous statements} are handled in
contract-based design.  The \emph{simultaneous statement} represents
analogue constraint of continuous-time waveform.  We will assume that
the constraint of a waveform is represented by an invariant. This
premise deserves an informal explanation.

Briefly stated, the analog kernel calculates quantities such that the
constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations.  It means that the number of simultaneous
statements should be equal to the number of unknowns
(\emph{quantities}).

For instance, when we reasoning about the conservation of energy law,
the \emph{simultaneous statements} constrain the values of
\emph{branch quantities} in a model.  In other words, the sum of
\arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) around
any closed path must be equal to zero and the sum of \through
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
(terminal) must be equal to zero.  To be more precise, the results of
summations above remains zero when a particular transformation is
applied to quantities.  In general, they are invariant.

@@

%=====================================================================
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}

In this section, we briefly describe the mapping between VHDL-AMS and
BON notation which is discussed in the points below.

@@

@@

%=====================================================================
\section{Conclusion}
\label{sec:conclusion}

The BON notation has originally been used for description of software
models. However, after certain elaboration it can be used for
description of hardware components. Practical utilization of the
approach in the project on mapping between VHDL-AMS and BON using
refinement technique, has demonstrated through low-pass filter example
and shown that simulation of component requires relatively little and
inexpensive efforts. Moreover, our article shown that the development
of the specification produced by the BON and translated into JML can
be applied as refinement procedure: from VHDL-AMS to informal BON,
next, from informal to more formal BON and finally, to JML
specification.  The BON support general software development and
should be adapted for concrete applications and language as
VHDL-AMS. For this purpose we have developed various VHDL-AMS related
types, operators and relations and we have defined appropriate
specifications for all of them. However, the behavior of
\texttt{break} statement and concurrency, for example, have not been
translated yet in the BON notation. We have started a new phase of
project named \emph{eBon} in which the issues mentioned above, and
creation of appropriate components and relations are main subject.
